Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.Rmd
take notes
clean comment

<!-- check the policy -->

<!-- problem: residual plot diagnostics -->

<!-- conventional test: too sensitive -->

<!-- background:  -->

<!-- week 1 -->

<!-- 1. residual plot for model diagnostics -->

<!-- a. residual is widely used -->

<!-- b. what are the types of residual plots -->

<!-- c. comparison -->

<!-- week 2 -->

<!-- 2. conventional test: F, BP -->

<!-- 3. visual test: lineup, theory -->

<!-- week 4 -->

<!-- desc of experiment: -->

<!-- 1. simulation setup -->

<!-- 2. experimental design -->

<!-- 3. result -->

<!-- week 3 -->

<!-- comparison of conventional tests: -->

<!-- last week -->

<!-- conclusion: -->

<!-- 1. too sensitive, visual test is needed/preferable -->

<!-- 2. visual test is infeasible in large scale (expensive) -->

<!-- 3. future work (role of computer vision) -->

@@

Heterskedasticity refers to the presence of nonconst error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \\ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.

@@

<!-- Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots. The debt on which type of residuals should be used is always present. While raw residuals are the most common computer regression software package output, by applying a scaling factor, the ability to reveal nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst1980regression]. As a two-dimensional representation of a model in a $p$-dimensional space, standard residual plots project data points onto the variable of the horizontal axis, which is a vector in $p$-dimensional space. Observations with the same projection will be treated as equivalent as they have the same abscissa. Consequently, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis. Hence, in practice, multiple standard residual plots with different horizontal axes will be examined [@cook1982residuals]. Overlapping data points is another general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed. Thus, for a relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aid to the eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.  -->

<!-- Other types of data plots that are often used in regression diagnostics include partial residual plots and probability plots. Partial residual plots are useful supplements to standard residual plots as they provide additional information on the extent of the non-linearity. Probability plots can be used to compare the sampling distribution of the residuals to the normal distribution for assessing the normality assumptions. -->

<!-- ## Hypothesis testing -->

<!-- Other than checking diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects. Depending on the alternative hypothesis that is focused on, a variety of tests can be applied. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. For testing non-linearity, one may apply the F-test to examine the significance of specific polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].  -->

<!-- As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers. @montgomery1982introduction suggested that based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies. Not to mention, it is almost impossible to have an exactly correctly specified model in reality. There is a well-known aphorism in statistics stated by George Box - "All models are wrong, but some are useful". This indicates proper hypothesis tests will always reject the null hypothesis as long as the sample size is large enough. The outcome "Not reject" can be interpreted as either "effect size is small" or "sample size is small". and the outcome "reject" doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction. But still, the effectiveness of statistical tests shall not be disrespected. Statistical tests have a chance to provide analysts with unique information. In situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots are needed to be checked, one will have no choice but to fall back on statistical tests if there exists any. A good regression diagnostic practice should be a balanced combination of both methods. -->

@@

@@

@@

@@

@@

>>>>

@@

Heterskedasticity refers to the presence of nonconst error variance in a regression model. It is mostly due to the strict but false assumptions on the variance-covariance matrix of the error term. The usual pattern of heterskedasticity on a residual plot is the inconsistent spread of the residuals at different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure \ref{fig:commondepartures}, or the "left-triangle" and "right-triganle" shape where the smallest variance occurs at the edges of the x-axis.

@@

@@

<!-- report 0 when the value is small -->

<!-- F-test -> model structrual test -->

<!-- F-test -> F -->

<!-- right justified  -->

<!-- test formula -> appendix -->

<!-- red col -> bold -->

@@

<!-- improve the first sentence -->

@@

<!-- use true exp lineup as example lineup -->

@@

<!-- remove the binomial one -->

@@

<!-- a sentence to discuss alpha  -->

### Power calculation

### 2022-09-20 paper_comparison/paper_comparison/paper_comparison.tex
clean comment

Heterskedasticity refers to the presence of nonconst error variance in a
regression model. It is mostly due to the strict but false assumptions
on the variance-covariance matrix of the error term. The usual pattern
of heterskedasticity on a residual plot is the inconsistent spread of
the residuals at different x values. Visually, it sometimes results in
the so-called ``butterfly'' shape as shown in the bottom-left panel of
Figure \textbackslash ref\{fig:commondepartures\}, or the
``left-triangle'' and ``right-triganle'' shape where the smallest
variance occurs at the edges of the x-axis.

@@

As pointed out by \citet{vanderplas2021statistical}, though Equation
(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
possible dependencies in the visual test due to repeated evaluations of
the same lineup. And it is inapplicable to visual test where subjects
are asked to select one or more ``most different'' plots from the
lineup. They summarized three common different scenarios in visual
inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
\(K\) lineups with different null plots but the same actual data plot
are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
subjects. Out of these three scenarios, Scenario 3 is the most common in
previous studies as it puts the least constraints on the experimental
design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Thus,
the visual p-value followed by the beta-binomial model is given as

@@

>>>>

Heterskedasticity refers to the presence of nonconst error variance in a
regression model. It is mostly due to the strict but false assumptions
on the variance-covariance matrix of the error term. The usual pattern
of heterskedasticity on a residual plot is the inconsistent spread of
the residuals at different x values. Visually, it sometimes results in
the so-called ``butterfly'' shape as shown in the bottom-left panel of
Figure \ref{fig:commondepartures}, or the ``left-triangle'' and
``right-triganle'' shape where the smallest variance occurs at the edges
of the x-axis.

@@

As pointed out by \citet{vanderplas2021statistical}, though Equation
(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
possible dependencies in the visual test due to repeated evaluations of
the same lineup. And it is inapplicable to visual test where subjects
are asked to select one or more ``most different'' plots from the
lineup. They summarized three common different scenarios in visual
inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
\(K\) lineups with different null plots but the same actual data plot
are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
subjects. Out of these three scenarios, Scenario 3 is the most common in
previous studies as it puts the least constraints on the experimental
design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Thus,
the visual p-value followed by the above model is given as

@@

\hypertarget{power-calculation}{%
\subsubsection{Power calculation}\label{power-calculation}}
